White Noise Test from Ordinal Patterns in the Entropy-Complexity Plane

### 2021-12-20 Reports/JOURNAL - Confidence Regions/IT-ConfidenceRegionsR2.tex
Version R2

\documentclass[alpha-refs]{wiley-article}

@@

\title{Time Series Analysis with Bandt \& Pompe Symbolization and a Test for White Noise in the Entropy-Complexity Plane}

@@

\begin{abstract}
This article serves two purposes.
Firstly, it surveys the \citeauthor{PermutationEntropyBandtPompe} methodology for the statistical community, stressing topics that are open for research.
Secondly, it contributes towards a better understanding of the statistical properties of that approach for time series analysis.
%
The \citeauthor{PermutationEntropyBandtPompe} methodology 
%has been used successfully in the analysis of time series. 
%It 
consists of computing
Information Theory descriptors of the histogram of ordinal patterns. 
Such descriptors lie in a 2D manifold: the Entropy-Complexity plane. 
This article provides the first proposal
of a test in the Entropy-Complexity plane for the
white noise hypothesis.
Our test is based on true white noise sequences obtained from physical devices.
The proposed methodology provides consistent results: 
it assesses sequences of true random samples as random (adequate test size), 
rejects correlated and contaminated sequences (sound test power), and
captures the randomness of generators previously analyzed in the literature.
\keywords{Random number generators, Bandt–Pompe approach, Entropy-complexity plane, Time Series, Information Theory}
\end{abstract}

@@

The Jensen-Shannon distance between $\bm h$ and the probability function $\bm u=\big(u(1), u(2), \dots, u(D!)\big)$ stems as an adequate measure of such a distance:
\begin{equation}
Q'(\bm{h}, \bm{u}) = \sum_{\ell=1}^{D!} \Big(h(\ell) \log\frac{h(\ell)}{u(\ell)} +
u(\ell) \log\frac{u(\ell)}{h(\ell)}
\Big).
\label{eq:JensenShannon}
\end{equation}
This has been the preferred distance function since its proposal by \citet{Lamberti2004}.
Most works consider the uniform distribution $\bm u = (1/D!, 1/D!, \dots, 1/D!)$ as the reference model.
In this case, \eqref{eq:JensenShannon} becomes
\begin{equation}
Q'(\bm{h}, \bm{u}) = \sum_{\ell=1}^{D!} h(\ell) \log h(\ell) - \frac{1}{D!} \sum_{\ell=1}^{D!} \log h(\ell).
\label{eq:JS-Uniform}
\end{equation}
The usual assumption in~\eqref{eq:JensenShannon} and~\eqref{eq:JS-Uniform} is $0\log 0=0$.

>>>>

% !TeX TS-program = pdflatex
\documentclass[alpha-refs]{wiley-article}

@@

\title{White Noise Test from Ordinal Patterns in the Entropy-Complexity Plane}

@@

\begin{abstract}
This article serves two purposes.
Firstly, it surveys the \citeauthor{PermutationEntropyBandtPompe} methodology for the statistical community, stressing topics that are open for research.
Secondly, it contributes towards a better understanding of the statistical properties of that approach for time series analysis.
%
The \citeauthor{PermutationEntropyBandtPompe} methodology 
%has been used successfully in the analysis of time series. 
%It 
consists of computing
Information Theory descriptors from the histogram of ordinal patterns. 
Such descriptors lie in a 2D manifold: the Entropy-Complexity plane. 
This article provides the first proposal
of a test in the Entropy-Complexity plane for the
white noise hypothesis.
Our test is based on true white noise sequences obtained from physical devices.
The proposed methodology provides consistent results: 
it assesses sequences of true random samples as random (adequate test size), 
rejects correlated and contaminated sequences (sound test power), and
captures the randomness of generators previously analyzed in the literature.
\keywords{Random number generators, Bandt–Pompe approach, Entropy-complexity plane, Ordinal Patterns, Time Series, Information Theory}
\end{abstract}

@@

The Jensen-Shannon distance between $\bm h$ and the probability function $\bm u=\big(u(1), u(2), \dots, u(D!)\big)$ stems as an adequate measure of such a distance:
\begin{equation}
Q'(\bm{h}, \bm{u}) = \sum_{\ell=1}^{D!} \Big(h(\ell) \log\frac{h(\ell)}{u(\ell)} +
u(\ell) \log\frac{u(\ell)}{h(\ell)}
\Big).
\label{eq:JensenShannon}
\end{equation}
This has been the preferred distance function since its proposal by \citet{Lamberti2004}.
Most works consider the uniform distribution $\bm u = (1/D!, 1/D!, \dots, 1/D!)$ as the reference model.
